---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

# Introduction

bisonR is an R package for running social network analyses in the BISoN framework. BISoN models consist of two main stages: 1) fitting an edge weight model, capturing uncertainty in social network edges; and 2) propagating this uncertainty through subsequent analyses such as regressions. In this short tutorial we'll cover how to fit edge weight models to a simulated dataset, and show how the fitted edge weight model can be used to run dyadic and nodal regression analysis, as well as non-random edge weight tests.

Before beginning, we'll load in the `bisonR` package using `library(bisonR)`, and we'll also bring in `dplyr` to help with data wrangling. If you don't already have `bisonR` installed, you can install it from Github using the following code:

```r
remotes::install_github("JHart96/bisonR")
```

```{r}
library(bisonR)
library(dplyr)
```

We will use the `simulate_bison_model()` function from `bisonR` to simulate some observation data. This dataframe is an example of the format that bisonR uses. Each row corresponds to an observation (such as an association within a sampling period of a dyad or a count of interactions between a dyad). Additional observation-level, dyad-level, or node-level factors can also be included.

```{r}
sim_data <- simulate_bison_model("binary", aggregated = FALSE)
df <- sim_data$df_sim
head(df)
```

# Edge Models

Depending on the type of data being analysed, different edge models should be used. See the `BISoN introduction` vignette for more details on this. In this example we're using binary data, where events can either occur (1) or not occur (0) in a particular sampling period. In our dataframe this corresponds to `event = 1` or `event = 0`.

The bisonR framework is fully Bayesian, and therefore specifying priors is a key part of any analysis. Priors can be tricky to set, and techniques for choosing good priors are outside the scope of this tutorial. You can see which priors need to be set for a particular model using the `get_default_priors()` function. These are only defaults and should almost always be changed.

```{r}
priors <- get_default_priors("binary")
priors
```

The priors object gives a list of priors on the parameters for the edge model: the edge weights, additional fixed effects, and random effect parameters. The `prior_check()` function can be used to plot the prior distributions, and we'll use this to check that we're happy with the priors:

```{r}
prior_check(priors, "binary")
```

The priors should encode our prior beliefs about likely values of the parameters. For demonstration purposes let's say we expect edge weights to be concentrated lower, towards zero. We can modify the priors on edge weight like this:

```{r}
priors$edge <- "normal(-1, 2.5)"
prior_check(priors, "binary")
```

Depending on the biology, this might be more like what we want to see. Now we've set our priors and we're happy with them, we can fit the edge model using the `bison_model()` function. This is one of the main functions of the `bisonR` package and has a lot of functionality. The main thing to worry about is how to define the edge weight model formula. The formula are designed to be familar to those who have used `lm`, `lme4`, `brms`, and other regression packages. The main difference here is that we need to include some measure of observation effort to get accurate estimates of edge weight uncertainty.

In bisonR, the left hand side of the `~` describes the sampling data, and uses the `(event | duration)` notation. In this notation, `event` corresponds to the name of the column in the dataframe that represents the measure of social events, such as 1 and 0 in the binary model, or frequencies 0, 1, 2, ... in the count model. If using an unaggregated dataframe, the `duration` corresponds to the durations of each observation. In our case this is fixed, but this will depend on the data at hand.

The right hand side of the `~` describes the predictors that are associated with social events. In the standard BISoN model this is primarily the edge weight that we use to build the network, but this can also include additional effects such as age, sex, or even observation-level factors such as location, time of day, or weather, to name a few. Including additional effects will change the interpretation of edge weights, but could be useful if interested in the social network once some factors have been controlled for. *Note* the edge weights don't need to be included, and you could use bisonR to estimate generative effects in a network. In this example we'll keep things simple and just use edge weight as the only predictor. This is done using the `dyad(node_1_id, node_2_id)` notation. `node_1_id` and `node_2_id` represent the nodes corresponding to each individual in the network, and need to be stored as factors in the dataframe. They don't need to be IDs, and could be individual names, as long as they're stored as factors. Now we understand the basics of the formula notation, we can now fit the edge weight model:

```{r}
fit_edge <- bison_model(
  (event | duration) ~ dyad(node_1_id, node_2_id), 
  data=df, 
  model_type="binary",
  priors=priors
)
```

Depending on your dataset this can take anywhere between a few tenths of a second and several hours. In particular if you have a large dataset and no observation-level predictors, it's probably a good idea to use the aggregated version of the model where all observations per dyad are collapsed to a single row. This will speed up model fitting considerably.

Once the model has fitted, we need to check that the MCMC algorithm has behaved correctly. Often if there's a major problem, the `bison_model()` function itself will have triggered warning messages above. But even if everything is silent, it's still worth checking the traceplots to check the MCMC chains have converged. The chains should be well-mixed and look something like a fuzzy caterpillar. We can check for this using the `plot_trace()` function:

```{r}
plot_trace(fit_edge, par_ids=2)
```

If we're satisfied that the MCMC algorithm has done its job properly, then it's time to see if the same is true for the model. One way to do this, among many others, is to check the predictions from the fitted model against the real data. This can be done using the `plot_predictions()` function. If the real data are within the ensemble of predictions from the model, then we have a little more faith that the model is capturing at least some important properties of the data. The statistical model implies a range of predictions that we can extract over multiple draws, the number of which can be set using the `num_draws` argument.

```{r}
plot_predictions(fit_edge, num_draws=20, type="density")
```

The predictions from the model are shown in blue and the real data are shown in black. Ideally the blue lines should be distributed around the black line, indicating the real data are among the possible predictions of the model. It's usually a good idea to run multiple predictive checks to ensure the model has captured various important characteristics of the data. Another type of check supported by `bisonR` is a comparison against point estimates of edge weights:

```{r}
plot_predictions(fit_edge, num_draws=20, type="point")
```

This plot shows BISoN estimates denoted by an interval, where the 90% probability interval is shown by the thin line, and the 50% probability interval is shown by the thicker blue line. We should see a roughly linear relationship between point estimates and BISoN estimates, though at the extremes the BISoN estimates are likely to be less extreme unless there are sufficient data. This is by design, as BISoN uses a full probabilistic model of social events to generate its estimates, and will be sceptical of extreme values without sufficient evidence.

Now we've conducted two basic diagnostic checks of the edge weight model, we can start to trust what it's telling us. To see a summary of edge weights and their credible intervals, we can use the `summary()` function:

```{r}
summary(fit_edge)
```

It can be hard to get an intuitive idea for what's going on looking only at a summary table, so this is where it can be useful to visualise the network too. The `plot_network()` function does this for BISoN edge weight models, where uncertainty is shown on edge weights. Uncertainty is visualised by showing the lower and upper bounds as overlapping edges in the network.

```{r}
plot_network(fit_edge, lwd=5)
```

# Non-random Edge Weights

Now that we have a fitted edge weight model we're happy with, we can move to downstream analyses. The first of these we'll consider is the non-random edge weight analysis, which is a Bayesian version of the Bejder et al. 1998 test for non-random association. In this analysis we compare the fitted edge weight model to a null version of the model where all edges have the same weight. We then use the `model_comparison` function to estimate the relative probabilities of the full edge weight model against the null model:

```{r}
fit_null <- bison_model(
  (event | duration) ~ 1, 
  data=df, 
  model_type="binary",
  priors=priors
)

model_comparison(list(non_random_model=fit_edge, random_model=fit_null))
```

This shows the relative probabilities of the models in the comparison. In our case the results are conclusive, the network seems to be non-random. It's likely that this is the case in almost all networks, but you might want to conduct this test anyway.

# Regression

One of the most common types of analysis is regression. To demonstrate how regressions can be conducted in bisonR, we'll first set up a dataframe. For this example let's use age difference as a predictor of edge weight.

```{r}
df_dyadic <- df %>%
  distinct(node_1_id, node_2_id, age_diff)

df_dyadic
```

bisonR uses the popular Bayesian regression framework `brms` to run regressions models inside a wrapper function `bison_brm()`. The core syntax and arguments to `bison_brm()` are the same as `brm`, with a few key differences. The main difference in the formula specification is that network variables are specified using the `bison(...)` keyword. This tells brms to use the bison edge weight model to maintain uncertainty in the analysis. In this analysis we want to use age difference as a predictor of edge weight, so we can run:

```{r}
fit_dyadic <- bison_brm (
  bison(edge_weight(node_1_id, node_2_id)) ~ age_diff,
  fit_edge,
  df_dyadic,
  num_draws=5, # Small sample size for demonstration purposes
  refresh=0
)
```

The default output of `bison_brm()` is a `brmsfit_multiple` object that can be analysed in much the same way as any other brms object. We can check the summary as follows:

```{r}
summary(fit_dyadic)
```

# Network Metrics

As well as the types of statistical analysis we've already covered, network metrics can also be extracted for use in other quantitative or qualitative analyses. This can be done with the `draw_network_metric_samples()` function for network-level metrics (also known as global metrics) or  `draw_node_metric_samples()` for node-level metrics. Let's try this out for the coefficient of variation (CV) in edge weights, also known as social differentiation:

```{r}
cv_samples <- extract_metric(fit_edge, "global_cv")
head(cv_samples)
```

We can visualise the posterior distribution of social differentiation by 

```{r}
plot(density(cv_samples))
```

The posterior samples can now be used descriptively, to compare networks, or in other downstream analyses.

# Conclusion

This was a very brief introduction to the `bisonR` package, and we hope it was useful. The R package has many more features than were shown here, and the package is under constant development. We welcome any and all feedback and criticism. We hope you find this package useful, happy social network analysis!



