---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

bisonR is an R package for running social network analyses in the BISoN framework. BISoN models consist of two main stages: 1) fitting an edge weight model, capturing uncertainty in social network edges; and 2) propagating this uncertainty through subsequent analyses such as regressions. In this short tutorial we'll cover how to fit edge weight models to a simulated dataset, and show how the fitted edge weight model can be used to run dyadic and nodal regression analysis.

Before beginning, we'll load in the `bisonR` package using `library(bisonR)`, and we'll also bring in `dplyr` to help with data wrangling. If you don't already have `bisonR` installed, you can install it from Github using the following code:

    remotes::install_github("JHart96/bisonR")

```{r}
library(bisonR)
library(dplyr)
```

We will also use the `simulate_edge_model()` function from `bisonR` to simulate some observation data. This dataframe is an example of the format that bisonR uses. Each row corresponds to an observation (such as an association within a sampling period of a dyad or a count of interactions between a dyad), but additional observation-level, dyad-level, or node-level factors can be included.

```{r}
sim_data <- simulate_edge_model("binary", aggregated = TRUE)
df <- sim_data$df_sim
head(df)
```
# Edge Models

Depending on the type of data being analysed, different edge models should be used. See the explanation of the bisonR framework here for more details on this. In this example we're using binary data, where events can either occur (1) or not occur (0) in a particular sampling period. In our dataframe this corresponds to `event = 1` or `event = 0`.

The bisonR framework is fully Bayesian, and therefore specifying priors are a key part of any analysis. Priors can be tricky to set, and techniques for choosing good priors are outside the scope of this tutorial, but see resources here for help on choosing priors in BISoN. We recommend retrieving the default priors from bisonR using the `get_default_priors()` function to get the correct set of default priors for your edge weight model.

```{r}
priors <- get_default_priors("binary")
priors
```

The priors object gives a list of priors on the parameters for the edge model: the edge weights, additional fixed effects, and random effect parameters. The `prior_check()` function can be used to plot the prior distributions, and we'll use this to check that we're happy with the priors:

```{r}
prior_check(priors, "binary")
```

Imagine we decide that the prior on edge weights isn't appropriate as we'd prefer a flatter prior for the edge weight parameter, we can change this using the following code:

```{r}
priors$edge <- "normal(-1, 1)"
prior_check(priors, "binary")
```

Depending on the biology, this might be more like what we want to see. Now we've set our priors and we're happy with them, we can fit the edge model using the `edge_model()` function. This is one of the main functions of the `bisonR` package and has a lot of functionality. The main thing to worry about is how to define the edge weight model formula. The formula are designed to be familar to those who have used `lm`, `lme4`, `brms`, and other regression packages. The main difference here is that we need to include some measure of sampling intensity to get accurate estimates of edge weight uncertainty.

In bisonR, the left hand side of the `~` describes the sampling data, and uses the `(event | duration)` notation. In this notation, `event` corresponds to the name of the column in the dataframe that represents the measure of social events, such as 1 and 0 in the binary model, or frequencies 0, 1, 2, ... in the count model. If using an unaggregated dataframe, the `duration` corresponds to the durations of each observation. In our case this is fixed, but this will depend on the data at hand.

The right hand side of the `~` describes the predictors that are associated with social events. In the standard BISoN model this is primarily the edge weight that we use to build the network, but this can also include additional effects such as age, sex, or even observation-level factors such as location, time of day, or weather, to name a few. Including additional effects will change the interpretation of edge weights, but could be useful if interested in the social network once some factors have been controlled for. *Note* the edge weights don't need to be included, and you could use bisonR to estimate generative effects in a network. In this example we'll keep things simple and just use edge weight as the only predictor. This is done using the `dyad(node_1_id, node_2_id)` notation. `node_1_id` and `node_2_id` represent the nodes corresponding to each individual in the network, and need to be stored as factors in the dataframe. They don't need to be IDs, and could be individual names, as long as they're stored as factors. Now we understand the basics of the formula notation, we can now fit the edge weight model:

```{r}
prior_predictive_check(
  (event | duration) ~ dyad(node_1_id, node_2_id),
  data=df, 
  model_type="binary",
  priors=priors
)
```

```{r}
fit_edge <- edge_model(
  (event | duration) ~ dyad(node_1_id, node_2_id), 
  data=df, 
  data_type="binary",
  priors=priors
)
```

Depending on your dataset this can take anywhere between a few tenths of a second and several hours. In particular if you have a large dataset and no observation-level predictors, it's probably a good idea to use the aggregated version of the model where all observations per dyad are collapsed to a single row. This will speed up model fitting considerably.

Once the model has fitted, we need to check that the MCMC algorithm has behaved correctly. Usually if there's a major problem, the `edge_model()` function itself will have triggered warning messages above. But even if everything is silent, it's still worth checking the traceplots to check the MCMC chains have converged. The chains should be well-mixed and look something like a fuzzy caterpillar. We can check for this using the `plot_trace()` function:

```{r}
plot_trace(fit_edge, par_ids=2)
```

If we're satisfied that the MCMC algorithm has done its job properly, then it's time to see if the same is true for the model. One way to do this, among many others, is to check the predictions from the fitted model against the real data. This can be done using the `plot_predictions()` function. If the real data are within the ensemble of predictions from the model, then we have a little more faith that the model is capturing at least some important properties of the data. The statistical model implies a range of predictions that we can extract over multiple draws, the number of which can be set using the `num_draws` argument.

```{r}
plot_predictions(fit_edge, num_draws=20, type="density")
```

The predictions from the model are shown in teal and the real data are shown in black. Ideally the blue lines should be distributed around the black line, indicating the real data are among the possible predictions of the model. Now we've conducted two basic diagnostic checks of the edge weight model, we can start to trust what it's telling us. To see a summary of edge weights and their credible intervals, we can use the `summary()` function:

```{r}
summary(fit_edge)
```

It can be hard to get an intuitive idea for what's going on looking only at a summary table, so this is where it can be useful to visualise the network too. The `plot_network()` function does this for BISoN edge weight models, where uncertainty is shown on edge weights. Uncertainty is visualised by showing the lower and upper bounds as overlapping edges in the network.

```{r}
plot_network(fit_edge, lwd=1)
```

Now that we have a fitted edge weight model we're happy with, we can start to look at downstream analyses. The first of these we'll consider is dyadic regression.


```{r}
fit_null <- edge_model(
  (event | duration) ~ 1, 
  data=df, 
  data_type="binary"
)

model_comparison(list(fit_edge, fit_null))
```

# Dyadic Regression

Dyadic regression is where dyadic network properties (usually edge weight) are regressed against dyad-level properties. In our case we'll be regressing edge weight against difference in age, to test the hypothesis that differences in age drive social connections. Just like edge weight models, we need to set priors, and this can be done in much the same way as last time:

```{r}
priors <- get_default_priors("dyadic_regression")
# priors$error <- "half-normal(2.5)"
priors
```

The dataframe for a dyadic regression will be different to the one used to fit an edge weight model. In this case we need a row for each dyad, and we need to index this using the same factor-typed node variables with the same levels as were used when fitting the edge weight model. In our example we have all the necessary data in the original dataframe, and we can extract it using dplyr. The `dyadic_regression()` function uses a similar formula syntax as the `edge_model` function. In our example we use the edge weight, described by `dyad(node_1_id, node_2_id)`, on the left hand side as the response variable. We use the `age_diff` variable from the dataframe as the single fixed effect, but additional fixed and random effects can also be included. *Note* by default the dyadic regression function uses multi-membership effects, denoted by the argument `mm`, to capture non-independence due to nodes in the regression. This is a vital feature for some analyses to prevent spurious conclusions, but for the sake of our simulated data we don't need to include this effect here.

```{r}
prior_predictive_check(
  dyad(node_1_id, node_2_id) ~ age_diff, 
  data=df_dyadic,
  options=list(edgemodel=fit_edge, mm=FALSE),
  priors=priors,
  model_type="dyadic_regression",
  plot_type="marginal"
)
```

```{r}
df_dyadic <- df %>%
  distinct(node_1_id, node_2_id, age_diff)

fit_dyadic <- dyadic_regression(
  dyad(node_1_id, node_2_id) ~ age_diff, 
  fit_edge, 
  df_dyadic, 
  mm=FALSE, 
  priors=priors
)
```

Like last time we want to check that our model has fit properly and makes good predictions. This can be done in exactly the same way as last time:

```{r}
plot_trace(fit_dyadic, par_ids=1)
```

```{r}
plot_predictions(fit_dyadic, num_draws=20, type="marginal", draw_data=FALSE)
```

Now we can examine the regression coefficients using `summary()`. These can be interpreted in mostly the same way as a standard regression, though categories will be treated differently. See here for more information on how categories are handled in `bisonR` regression functions.

```{r}
summary(fit_dyadic)
```

Now we've seen how to fit a dyadic regression, the next common type of analysis is nodal regression.

# Nodal Regression

Nodal regression is where node-level network properties, such as centrality, are regressed against node-level properties, such as age. In the same spirit as last time we need to prepare a matrix where each row corresponds to a node. In our case we can extract this from the original matrix using `dplyr`:

```{r}
df_1 <- df %>% 
  select(node_id=node_1_id, age=age_1)
df_2 <- df %>%
  select(node_id=node_2_id, age=age_2)
df_nodal <- bind_rows(df_1, df_2) %>%
  distinct(node_id, age) %>%
  mutate(age_std=(age - mean(df_nodal$age))/sd(df_nodal$age))
# df_nodal$age_std <- df_nodal$age - mean(df_nodal$age)
df_nodal
```

Priors can be set in the exact same way as for the `dyadic_regression()` function:

```{r}
priors <- get_default_priors("nodal_regression")
priors$fixed <- "normal(0, 1)"
priors
```

```{r}
prior_check(priors, model_type="nodal_regression")
```

The formula syntax for nodal regression is very similar to that of the dyadic regression, but because this analysis is at the level of the node, not the dyad, we no longer need a `dyad(node_1_id, node_2_id)` term. Instead we want to use a node-level network metric to characterise node positions within the network, while maintaining uncertainty from the edge weights. `bisonR` can handle all of this automatically, you simply need to specify a network metric to use. In our case we'll use the classic measure of centrality `strength()`, but many others are available. See the documentation here for further details.

```{r}
prior_predictive_check(
  strength(node_id) ~ age_std,
  data=df_nodal,
  model_type="nodal_regression",
  options=list(edgemodel=fit_edge),
  plot_type="marginal"
)
```

```{r}
fit_nodal <- nodal_regression(
  strength(node_id) ~ age_std, 
  fit_edge, 
  df_nodal, 
  priors=priors
)
```

Once the model has been fitted, we need to employ the same diagnostic checks as last time to ensure the model is reliable. This may seem a little tedious, and it won't always go smoothly, but it's a vital part of model fitting that is essential for reliable inferences, even in frequentist statistics.

```{r}
plot_trace(fit_nodal, par_ids=1)
```


```{r}
plot_predictions(fit_nodal, num_draws=20, type="marginal")
```

Just like the previous example, we can now summarise the regression coefficients using the `summary()` function:

```{r}
summary(fit_nodal)
```

# Network metrics

As well as the types of statistical analysis we've already covered, network metrics can also be extracted for use in other quantitative or qualitative analyses.

```{r}
draw_network_metric_samples(fit_edge, "social_differentiation")
```

# Conclusion

This was a very brief introduction to the `bisonR` package, and we hope it was useful. The R package has many more features than were shown here, and the package is under constant development. We welcome any and all feedback and criticism. We hope you find this package useful, happy social network analysis!



