---
title: "Edge mixture models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Edge mixture models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{igraph, brms}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=60),
  tidy=TRUE
)
```

# Introduction

Edge mixture models can be used to detect different types of social connections by looking for clustering in edge weights. bisonR provides support for edge mixture models with the `bison_mixture` function. There are two key outputs from a bisonR edge mixture model. First it generates a posterior probability distribution over the number of components (clusters, interpreted associal connection types), from 1 to a maximum $K$, over the entire network. Secondly, for each edge it calculates the probability that each dyad belongs to each possible component. To show how this all works, first we'll load in bisonR and simulate some data:

```{r}
library(bisonR)
library(dplyr)
library(igraph)
library(brms)
library(mclust)
```

In this example we simulate data from a binary edge model with 2 underlying clusters (also called components), with dyads belong to either cluster with equal probability:

```{r}
sim_data <- simulate_bison_model_mixture("binary", num_components = 2, component_weights = c(0.5, 0.5))
df <- sim_data$df_sim
head(df)
```

Since the data are already aggregated (e.g. a single row for each dyad), we can use the binary conjugate model for speed:

```{r}
fit_edge <- bison_model(
  (event | duration) ~ dyad(node_1_id, node_2_id), 
  data=df, 
  model_type="binary_conjugate"
)
```

We won't conduct the usual checks for the sake of this tutorial, but at this point we'd recommend conducting diagnostic checks to ensure model fit. Once the model has fit well, we can use the `bison_mixture()` function to fit the edge mixture model:

```{r}
# verbose=FALSE for tutorial purposes
fit_mixture <- bison_mixture(fit_edge, num_components=5, verbose=FALSE)
summary(fit_mixture)
```

There's quite a bit going on in the summary, so let's break it down. The top section shows the model that fits best and the probability assigned to that model. The second section shows the probabilities of the models corresponding to each number of components ($K = 1, 2, 3, ...$). You'll notice the model with the highest probability is the same as the best fitting model from the section above. The final section shows the edge-level probabilities of component membership for the corresponding dyad. This can be interpreted as the probability a dyad belongs to a given component/cluster/social connection type.

The information shown in the summary above can also be accessed using the functions `get_network_component_probabilities()` and `get_edge_component_probabilities()`. These two functions make it possible to access the mixture model output programmatically for downstream analysis or plotting. 

To get network-level probabilities of the number of components, we can use `get_network_component_probabilities()` function:

```{r}
get_network_component_probabilities(fit_mixture)
```
To use probabilities over component membership for a given number of components, we can use the `get_edge_component_probabilities()` function, where `3` is the number of components we assume to exist in the network:

```{r}
get_edge_component_probabilities(fit_mixture, 3)
```

It's often useful to know where the mixture components are location, so we can get their posterior means using the `get_component_means()` function for a given number of components. Here we'll calculate the posterior means of the components for the mixture model with 3 components:

```{r}
get_component_means(fit_mixture, 3)
```

# Using components in downstream analysis

Now say that we're interested in the number of strong vs weak partners. Specifically we want to know if the strength of connections only to strong partners predicts a node-level trait. This is a non-standard analysis so will require some manual data wrangling to fit the right kind of model.

For demonstration purposes, we'll create a dataframe that might be used for this kind of analysis:

```{r}
df_nodal <- data.frame(node_id=as.factor(levels(df$node_1_id)), trait=rnorm(15))
df_nodal
```
This dataframe will provide the information about traits that we need for the analysis. Next, we need to calculate strength using only individuals categorised as strong according to the mixture model. Since both component membership (strong or weak) *and* edge weights are probabilistic, we have to sample from both posteriors as we build a new posterior of mixture-based strength:

```{r}
# Set number of nodes and number of posterior samples
num_nodes <- fit_edge$num_nodes
num_draws <- 5 # Keep short for demo purposes.

# Create a list of igraph networks from edgemodel to represent network posterior
nets <- bison_to_igraph(fit_edge, num_draws)

# Create an empty matrix to hold strengths of top mixture component
mix_strengths <- matrix(0, num_draws, num_nodes)

# Create an empty list for imputed versions of the dataframe
imputed_data <- list()

# Loop through each posterior sample and calculate strength of top mixture.
for (i in 1:num_draws) {
  # Calculate edge components (1 if strong, 0 if weak)
  edge_components <- 1 * (fit_mixture$edge_component_samples[i, 2, ] == 2)
  
  # If the edge is strong, don't change edge weight, but if it's weak then set to zero.
  E(nets[[i]])$weight <- E(nets[[i]])$weight * edge_components
  
  # Change the metric values of the imputed data to be mixture-based strength
  imputed_data[[i]] <- df_nodal
  imputed_data[[i]]$mix_strength <- strength(nets[[i]])
}
imputed_data[[2]]
```

Now we have a set of imputed dataframes containing the posteriors of mixture-based node strength, we can fit the model using brm_multiple. 

```{r}
fit_brm <- brm_multiple(trait ~ mix_strength, imputed_data, silent=2, refresh=0)
summary(fit_brm)
```

This was a short tutorial on how to use `bison_mixture()` to fit mixture models in bisonR. If you have any questions please get in touch.

