---
title: "Network metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

Network metrics can be extracted using the function `extract_metric(bison_model, metric_name)`. The `metric_name` variable is made up of two parts: the network feature and the network metric. For example, `node_eigen` will calculate the eigenvector centrality of each node.

The following feature/metric pairs are currently supported in bisonR:

* `edge` - Edge-level properties
  * `weight` - Edge weight
* `node` - Node-level properties
  * `degree[t]` - Node degree (where t is the binary threshold for an edge)
  * `strength` - Node strength
  * `eigen` - Node eigenvector centrality
  * `betweenness` - Node betweenness
  * `closeness` - Node closeness
  * `clustering` - Node-level clustering coefficient
* `global` - Network-level properties
  * `density` - Weighted density
  * `cv` - Coefficient of variation of edge weights - also known as social differentiation
  * `std` - Standard deviation of edge weights
  * `diameter` - Diameter of the network
  * `clustering` - Clustering coefficient of the network - also known as transitivity
  
# Examples

See below for examples of how to generate the network metrics listed above.
  
```{r}
library(bisonR)
library(dplyr)
```
  
```{r}
sim_data <- simulate_bison_model("binary", aggregated = TRUE)
df <- sim_data$df_sim

fit_edge <- bison_model(
  (event | duration) ~ dyad(node_1_id, node_2_id),
  data=df,
  model_type="binary_conjugate",
  priors=get_default_priors("binary_conjugate")
)
```

## Edge

### Edge weight

```{r}
metrics <- extract_metric(fit_edge, "edge_weight", num_draws=10)
metrics[1:6, 1:5] # Preview first few samples and edges
```

## Node

### Node degree

```{r}
metrics <- extract_metric(fit_edge, "node_degree[0.2]")
metrics[1:6, 1:5] # Preview first few samples and nodes
```

### Node strength

```{r}
metrics <- extract_metric(fit_edge, "node_strength", num_draws=10)
metrics[1:6, 1:5] # Preview first few samples and nodes
```

### Node eigenvector centrality

```{r}
metrics <- extract_metric(fit_edge, "node_eigen")
metrics[1:6, 1:5] # Preview first few samples and nodes
```

### Node betweenness

```{r}
metrics <- extract_metric(fit_edge, "node_betweenness")
metrics[1:6, 1:5] # Preview first few samples and nodes
```

### Node closeness

```{r}
metrics <- extract_metric(fit_edge, "node_closeness")
metrics[1:6, 1:5] # Preview first few samples and nodes
```

### Node-level clustering coefficient

```{r}
metrics <- extract_metric(fit_edge, "node_clustering[0.2]")
metrics[1:6, 1:5] # Preview first few samples and nodes
```

## Global

### Global coefficient of variation

```{r}
metrics <- extract_metric(fit_edge, "global_cv")
metrics[1:6] # Preview first few samples
```

### Global density

```{r}
metrics <- extract_metric(fit_edge, "global_density")
metrics[1:6] # Preview first few samples
```

### Global standard deviation of edge weights

```{r}
metrics <- extract_metric(fit_edge, "global_std")
metrics[1:6] # Preview first few samples
```

### Global diameter

```{r}
metrics <- extract_metric(fit_edge, "global_diameter")
metrics[1:6] # Preview first few samples
```

### Global clustering coefficient

```{r}
metrics <- extract_metric(fit_edge, "global_clustering[0.2]")
metrics[1:6] # Preview first few samples
```
